---
title: "Statistical Analysis of Chronodel Test"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
date: "2026-04-09"
---

# Chronodel Data Analysis

The project that I am working on is to analyze the data of the **Chronodel test**, which is a test that measures the ability to perform a task in a delimation of time of 30 seconds. The data is collected from a sample of participants who have taken the test (n = 43). The goal of this analysis is to experiment the test Chronodel understand the distribution of the results and to compare the results of the Chronodel test with other tests that measure similar abilities, by others, i mean, Day Of the Week Backward and (DOTWB) and Month Of The Year Backward (MOTYB).

At the beginning of the analysis, we wanted to calculate the **sensibility**, **specificity**, **positive predictive** value and **negative predictive value** of the Chronodel test, but at the reception of the results, we saw, that there are a variability in the results of the Chronodel test, which is really low, and that the majority of the patients have a result of 1 on the Chronodel test, which means that they succeed the test, because we expected to have a more variability in the results, and also because we expected to have a lower success rate for the Chronodel test, so we decided to change our analysis plan and to focus on pourcentage of success for the three tests (DOTWB, MOTYB) to evaluate the coherence between the results and at the end, the distribution of the results for the three tests, to evaluate the variability and the ceiling effect.

A small number of missing values were observed across variables. As these were limited and not overlapping across tests, analyses were conducted using available data, with missing values excluded where necessary.

# inclusion of the data

The data cleaning data has been made earlier, in visual studio code, and the data is **ready to be analyzed**, so we will start by loading the data in Rstudio.

```{r}
Data <- read.csv("Stat_Chronodel_Rstudio_test.csv")
```

# Total subjects

To do the kappa test, we need to have complete cases for the variables "jour_reussite" and "mois_reussi", so we will check how many complete cases we have for these two variables.

```{r}
sum(complete.cases(Data[, c("jour_reussite", "mois_reussi")]))
```

We have a total of **36 completes cases** for these two variables, which is enough to do the Kappa test.

# Kappa test

```{r}
library(irr)
kappa2(Data[, c("jour_reussite", "mois_reussi")])
```

## Kappa test results

As result of the Kappa test, we have a **kappa value** of **0.0859**, which is a really low value of agreement between the DOTWB and MOTYB tests. This suggest that the observed agreement is only slightly above what would be expected by chance. This low Kappa value is likely explained by the highly unbalanced distribution of the data, with a majority of successful outcomes in the day-based test (ceiling effect), as well as the presence of discordant cases.

The p-value associated with the Kappa statistic was **0.204**, which is above the conventional significance threshold of **0.05**. Therefore, we fail to reject the null hypothesis (H0: κ = 0), meaning that there is no statistically significant agreement beyond chance between the two tests."

# Cross table 

We will create a cross table to estimate the coherence between the results of the two tests (DOTWB and MOTYB).

```{r}
table(Data$jour_reussite, Data$mois_reussi, useNA = "no")
```

## DOTWB vs MOTYB Concordance
 
 From the cross table, we observe: 
 
- 1 subject failed both DOTWB and MOTYB  
- 0 subjects failed DOTWB but succeeded MOTYB  
- 13 subjects succeeded DOTWB but failed MOTYB  
- 22 subjects succeeded both tests  

This results in 63.9% concordance (22+1/36) and **36.1% discordance** (13/36), explaining the low Kappa value.

# success rate

We will do now, the analysis of the success rate for the three tests, to evaluate the coherence between the results of the three tests, and to evaluate the ceiling effect for each test.

```{r}
taux_jour <- mean(Data$jour_reussite, na.rm = TRUE)
taux_mois <- mean(Data$mois_reussi, na.rm = TRUE)
taux_chrono <- mean(Data$chrono_reussi, na.rm = TRUE)

taux <- c(taux_jour, taux_mois, taux_chrono)

barplot(taux,
        names.arg = c("DOTWB", "MOTYB", "Chronodel"),
        ylim = c(0,1.1),
        ylab = ("success rate"),
        main = ("success rate for the three tests"),
        col = c("red", "blue", "green"))
text(x=1, y=taux_jour+0.05, labels=paste0(round(taux_jour*100, 2), "%"))
text(x=2, y=taux_mois+0.05, labels=paste0(round(taux_mois*100, 2), "%"))
text(x=3, y=taux_chrono+0.05, labels=paste0(round(taux_chrono*100, 2), "%"))
```

For the result of the three tests, we have a success rate of 97.3% for the DOTWB test, which is really high, and might suggest a ceiling effect, we have a success rate of 58.97% for the MOTYB test, which is more balanced, and that doesn't suggest a ceiling effect, and at the end, we have a success rate of 100% for the Chronodel test, which is really high, and that too as well a ceiling effect.

# Distribution 

For the final graph, we will plot the distribution of the results for the three tests, to evaluate the variability and the ceiling effect for each test. 

```{r}
par(mfrow=c(1,3))
barplot(table(Data$jour_reussite, useNA = "no")/nrow(Data),
        main="Distribution DOTWB", ylim=c(0,1.1), col=c("red", "green"), names.arg = c("0", "1"))
text(x=1, y=table(Data$jour_reussite, useNA = "no")[1]/nrow(Data)+0.05, labels=paste0(round(table(Data$jour_reussite, useNA = "no")[1]/nrow(Data)*100, 2), "%"))
text(x=2, y=table(Data$jour_reussite, useNA = "no")[2]/nrow(Data)+0.05, labels=paste0(round(table(Data$jour_reussite, useNA = "no")[2]/nrow(Data)*100, 2), "%"))
barplot(table(Data$mois_reussi, useNA = "no")/nrow(Data),
        main="Distribution MOTYB", ylim=c(0,1.1), col=c("red", "green"), names.arg = c("0", "1"))
text(x=1, y=table(Data$mois_reussi, useNA = "no")[1]/nrow(Data)+0.05, labels=paste0(round(table(Data$mois_reussi, useNA = "no")[1]/nrow(Data)*100, 2), "%"))
text(x=2, y=table(Data$mois_reussi, useNA = "no")[2]/nrow(Data)+0.05, labels=paste0(round(table(Data$mois_reussi, useNA = "no")[2]/nrow(Data)*100, 2), "%"))
barplot(c(0, mean(Data$chrono_reussi, na.rm = TRUE)),
        main="Distribution Chronodel", ylim=c(0,1.1), col=c("red", "green"), names.arg = c("0", "1"))
text(x=1, y=0.05, labels="0%")
text(x=2, y=mean(Data$chrono_reussi, na.rm = TRUE)+0.05, labels=paste0(round(mean(Data$chrono_reussi, na.rm = TRUE)*100, 2), "%"))
par(mfrow=c(1,1))
```

The distribution of the results for the DOTWB test is really unbalanced, with a majority of successful outcomes, which suggest **clearly** a ceiling effect. The distribution of the results for the MOTYB test is more balance, with 40% of "no" and 57.5% of "yes", which doesn't suggest a ceiling effect. The distribution of the results for the Chronodel test is really unbalanced, with all the subjects having a successful outcome, which suggest **strongly** a ceiling effect. This ceiling effect can be explained by the fact that the tests are not challenging enough for the subjects, and that the subjects are able to perform the task without any difficulty, which is not what we expected, and that is why we have a low variability in the results, and a high success rate for the three tests.

